---
title: "Ownership Database"
output: html_document
date: "2024-05-30"
---

In this notebook I create a "propiedad" (ownership) database with the data of only tenure regimes of each municipality to make it easier to get this data. "Ownership" includes both ownership by purchase and by inheritance. I also create a "target21" and "target11" database which includes all the individuals in the target population.

This takes long to run, specially loading the Census Microdata. The databases are uploaded on GitHub, so it's not needed to run everything to examine them.

# Packages

```{r}
library(tidyverse)
library(ggpubr)
library(ggplot2)
library(patchwork)
library(readxl)
library(haven)
library(flextable)
library(openxlsx)

library(extrafont)

# Importar fuentes del sistema operativo
#font_import(prompt = FALSE)

# Cargar fuentes
loadfonts(device = "win")

```

# Creating "Onwnership" database

### Create cusec and add names

```{r}

# Microdata 2011
microdatos2011 <- read_sav("Microdatos2011/Personas/poblacion_2011_v2.sav")

propiedad <- microdatos2011 |> 
  select(CPRO, CMUN)

# Create cusec codes of rural municipalities and add names
cusec_991 <- sprintf("%05d", seq(1991, 52991, by = 1000))
cusec_992 <- sprintf("%05d", seq(1992, 52992, by = 1000))
cusec_993 <- sprintf("%05d", seq(1993, 52993, by = 1000))
cusec_994 <- sprintf("%05d", seq(1994, 52994, by = 1000))
cusec <- c(cusec_991, cusec_992, cusec_993, cusec_994)

MUN_LITERAL <- 
  ifelse(substr(cusec, 4, 5) == "91","<=2000",
  ifelse(substr(cusec, 4, 5) == "92", "2001 <= 5000",
  ifelse(substr(cusec, 4, 5) == "93", "5001 <= 10000",
  ifelse(substr(cusec, 4, 5) == "94", "10001 <= 20000","Otros"))))

df_rural <- data.frame(MUN_LITERAL, cusec)

# Get the names of all municipalities and provinces
locations <- read_excel("vivienda turística/exp_viv_turistica_tabla5_FEB2021.xlsx", sheet = 3, na = "Dato protegido por secreto estadístico.")

nombres_municipios <- locations |> 
  select(MUN_LITERAL, MUN) |> 
  rename(cusec = MUN) |> 
  distinct(cusec, .keep_all = TRUE)

nombres_provincias <- locations |> 
  select(PROV_LITERAL, PROV) |> 
  rename(CPRO = PROV) |> 
  distinct(CPRO, .keep_all = TRUE)

nombres_municipios <- rbind(nombres_municipios, df_rural)

# Create names CCAA
ccaas <- data.frame(
  CodigoProvincia = c("01", "02", "03", "04", "33", "05", "06", "08", "09", "10", "11", "39", "12", "13", "14", "16", "17", "18", "19", "20", "21", "22", "23", "07", "24", "15", "35", "25", "26", "27", "28", "29", "30", "31", "32", "34", "36", "37", "38", "39", "40", "41", "42", "43", "44", "45", "46", "47", "48", "49", "50", "51", "52"),
  CCAA = c("País Vasco", "Castilla-La Mancha", "Comunidad Valenciana", "Andalucía", "Asturias", "Castilla y León", "Extremadura", "Cataluña", "Castilla y León", "Extremadura", "Andalucía", "Cantabria", "Comunidad Valenciana", "Castilla-La Mancha", "Andalucía", "Castilla-La Mancha", "Cataluña", "Andalucía", "Castilla-La Mancha", "País Vasco", "Andalucía", "Aragón", "Andalucía", "Islas Baleares", "Castilla y León", "Galicia", "Canarias", "Cataluña", "La Rioja", "Galicia", "Comunidad de Madrid", "Andalucía", "Comunidad de Murcia", "Comunidad Foral de Navarra", "Galicia", "Castilla y León", "Galicia", "Castilla y León", "Canarias", "Cantabria", "Castilla y León", "Andalucía", "Castilla y León", "Cataluña", "Aragón", "Castilla-La Mancha", "Comunidad Valenciana", "Castilla y León", "País Vasco", "Castilla y León", "Aragón", "Ceuta", "Melilla"))

# Add municipality and province names to each cusec
propiedad <- propiedad |> 
  mutate(
    CMUN = str_pad(CMUN, width = 3, side = "left", pad = "0"),
    CPRO = str_pad(CPRO, width = 2, side = "left", pad = "0")) |> 
   mutate(  
    CPRO = trimws(CPRO),
    CMUN = trimws(CMUN),
    cusec = paste0(CPRO, CMUN)) |> 
  left_join(nombres_municipios, by = "cusec") |> 
  left_join(nombres_provincias, by = "CPRO") |> 
  left_join(ccaas, by = c("CPRO" = "CodigoProvincia")) |> 
  distinct(cusec, .keep_all = T)

```

### Rural municipalities

Aquí se solucionan unas pequeñas discrepancias entre los municipios del Censo del 2011 y 2021. Se deben a que los municipios en 2011 entre 10000 a 20000 habitantes están agrupados dentro del código "994", pero en 2021 éstos tienen su nombre completo. Además, algunos municipios se encuentran en el límite de haber pasado una de las barreras de estos códigos entre ambos años y provoca algunos pequeños desajustes.

Datos de población: Padrón municipal e INE. <https://www.ine.es/dynt3/inebase/es/index.htm?padre=525>

```{r}

# Load data 2021
load("Microdatos2021/Personas/R/CensoPersonas_2021.RData")

microdatos2021 <- Microdatos

# Get all 2021 municipalities
mun21 <- microdatos2021 |> 
  select(CPRO, CMUN) |>    
  mutate(CPRO = trimws(CPRO), CMUN = trimws(CMUN), cusec = paste0(CPRO, CMUN)) |> 
  distinct(cusec)

# Get all 2011 municipalities
mun11 <- microdatos2011 |> 
  select(CPRO, CMUN) |> 
   mutate(
    CMUN = str_pad(CMUN, width = 3, side = "left", pad = "0"),
    CPRO = str_pad(CPRO, width = 2, side = "left", pad = "0")) |>
   mutate(CPRO = trimws(CPRO), CMUN = trimws(CMUN), cusec = paste0(CPRO, CMUN)) |> 
  distinct(cusec)

# These are the municipalities lost from 2021 to 2011. 
# Municipalities from 10000 to 20000 inhabitants
municipios994 <- mun21 |> 
  anti_join(mun11, by = "cusec") |> 
  mutate(cusec_994 = cusec)

mun_994 <- municipios994 |> 
  mutate(cusec = paste0(substr(cusec, 1, 2), "994")) 

# Add code to municipalities according to population 2011
poblacion_mun11 <- read_xls("Población/pobmun11.xls")

poblacion_mun11 <- poblacion_mun11 |> 
  setNames(poblacion_mun11[1, ]) |> 
      slice(-1) |> 
    select(-c(MUJERES, VARONES)) |> 
  rename(Poblacion2011 = `AMBOS SEXOS`) |> 
    mutate(CPRO = trimws(CPRO), CMUN = trimws(CMUN), cusec = paste0(CPRO, CMUN)) 

cusec_rural_2011 <- poblacion_mun11 |> 
  mutate(Poblacion2011 = as.numeric(Poblacion2011)) |> 
  filter(Poblacion2011 < 20000) |> 
  mutate(cusec_original = cusec,
    cusec = case_when(
      Poblacion2011 < 2000 ~ paste0(substr(cusec_original, 1, 2), "991"),
      Poblacion2011 < 5000 ~ paste0(substr(cusec_original, 1, 2), "992"),
      Poblacion2011 < 10000 ~ paste0(substr(cusec_original, 1, 2), "993"),
      Poblacion2011 < 20000 ~ paste0(substr(cusec_original, 1, 2), "994"),
      TRUE ~ cusec
    )
  )

cusec_rural_2011 <- cusec_rural_2011 |> 
  distinct(cusec_original, .keep_all = TRUE) |> 
  select(cusec_original, cusec)


# Assign code to municipalities according to population 2021
poblacion_mun21 <- read_xlsx("Población/pobmun21.xlsx")

poblacion_mun21 <- poblacion_mun21 |> 
  setNames(poblacion_mun21[1, ]) |> 
      slice(-1) |> 
    select(-c(MUJERES, HOMBRES)) |> 
  rename(Poblacion2021 = `POB21`) |> 
    mutate(CPRO = trimws(CPRO), CMUN = trimws(CMUN), cusec = paste0(CPRO, CMUN)) 

cusec_rural_2021 <- poblacion_mun21 |> 
  mutate(Poblacion2021 = as.numeric(Poblacion2021)) |> 
  filter(Poblacion2021 < 20000) |> 
  mutate(cusec_original = cusec,
    cusec = case_when(
      Poblacion2021 < 2000 ~ paste0(substr(cusec_original, 1, 2), "991"),
      Poblacion2021 < 5000 ~ paste0(substr(cusec_original, 1, 2), "992"),
      Poblacion2021 < 10000 ~ paste0(substr(cusec_original, 1, 2), "993"),
      Poblacion2021 < 20000 ~ paste0(substr(cusec_original, 1, 2), "994"),
      TRUE ~ cusec
    )
  )

cusec_rural_2021 <- cusec_rural_2021 |> 
  distinct(cusec_original, .keep_all = TRUE) |> 
  select(cusec_original, cusec)

```

## Filtrar población objetivo

### 2011

```{r}

# Select variables 2011
compra11 <- microdatos2011 |> 
  select(CPRO, CMUN, EDAD, ANAC, NORDEN, ANORES, TENEN, OPA_NORDEN, TIPOPER, ESTHOG, EDADPAD, EDADMAD) 

# Filter target population
compra11 <- compra11 |> 
  filter(EDAD %in% c(18:35), # Young people
         TIPOPER != "H", # That aren't "child" in household
         is.na(EDADPAD), # Don't live with father
         is.na(EDADMAD)) |> # Don't live with mother 
    select(-EDADPAD, -EDADMAD) |> 
  filter(ESTHOG %in% c(01, 02, 05, 07, 08, 10, 11)) # Isn't a home with a only person over 65+ or children over 25

compra11 <- compra11 |> 
  mutate(mayor = ifelse(ANORES - ANAC > 18, TRUE, FALSE)) |> 
  filter(mayor == TRUE) # Of legal age when last moved
  
compra11 <- compra11 |> 
  filter(NORDEN %in% c(1, 2)) # Order 1 or 2 in household

# Recode tenure regime
compra11 <- compra11 |>  
  mutate(
    TENEN_r = case_when(
      TENEN == 1 ~ "Propiedad por compra, pagada",
      TENEN == 2 ~ "Propiedad por compra, pagos pendientes",
      TENEN == 3 ~ "Propiedad por herencia o donación",
      TENEN == 4 ~ "Alquiler",
      TENEN == 5 ~ "Cedida o a bajo precio",
      TENEN == 6 ~ "Otro")) |> 
  mutate(
    TENEN_VIV = case_when(
      TENEN_r %in% c("Propiedad por compra, pagada", "Propiedad por compra, pagos pendientes") ~ "Compra",
      TENEN_r == "Propiedad por herencia o donación" ~ "Herencia o donación",
      TENEN_r == "Alquiler" ~ "Alquiler",
      TENEN_r %in% c("Cedida o a bajo precio", "Otro") ~ "Otro"))

# Add cusec
compra11 <- compra11 |>  
  mutate(
    CMUN = str_pad(CMUN, width = 3, side = "left", pad = "0"),
    CPRO = str_pad(CPRO, width = 2, side = "left", pad = "0")) |> 
   mutate(  
    CPRO = trimws(CPRO),
    CMUN = trimws(CMUN),
    cusec = paste0(CPRO, CMUN))

compra11 <- compra11 |> 
  left_join(nombres_municipios, by = "cusec") |> 
  left_join(nombres_provincias, by = "CPRO") |> 
  left_join(ccaas, by = c("CPRO" = "CodigoProvincia"))


```

### 2021

```{r}
# 2021
compra21 <- microdatos2021 |> 
  select(CPRO, CMUN, VAREDAD, ANAC, TENEN_VIV, NORDEN, RESI_NACIM, VARANORES, VARANOM, CMUN_ANT, CMUN_UNANO, RESI_UNANO, CMUN_DANO, RESI_DANO, LTRAB, CMUN_TRAB, LEST, CMUN_EST, NORDEN_CON, NORDEN_OPA, TIPOPER, SEXO_MAD, SEXO_PAD, ESTRUC_HOG) 

compra21 <- compra21 |> 
  filter(VAREDAD %in% c(18:35), # Young 
         TIPOPER != "H", # Not "child"
         SEXO_PAD == "N", # Don't live with father
         SEXO_MAD == "N") |>  # Don't live with mother
  mutate(ESTRUC_HOG = as.numeric(ESTRUC_HOG)) |> 
  filter(ESTRUC_HOG %in% c(01, 02, 05, 07, 08, 10, 11)) # Not an only person over 65 or children over 25

# Last move being over 18 yo (risking adding some irrelevant cases)
compra21 <- compra21 |> 
  mutate(
    VARANORES_r = case_when(
      VARANORES == "0005" ~ as.numeric(2015),
      VARANORES == "0004" ~ as.numeric(2010),
      VARANORES %in% c("9999", "0003") ~ as.numeric(2000), 
      VARANORES == " " ~ NA,
      TRUE ~ as.numeric(VARANORES)),
      ANAC = as.numeric(ANAC),
      mayor = ifelse(VARANORES_r - ANAC > 18, TRUE, FALSE)) |> 
  filter(mayor == TRUE)

compra21 <- compra21 |> 
  filter(NORDEN %in% c("01", "02")) # Order 1 or 2 in the household

# Recode tenure regime
compra21 <- compra21 |> 
    mutate(
    TENEN_VIV = case_when(
      TENEN_VIV == 2 ~ "Propiedad",
      TENEN_VIV == 3 ~ "Alquiler",
      TENEN_VIV == 4 ~ "Otro")) 

# Add cusec
compra21 <- compra21 |> 
  mutate(CPRO = trimws(CPRO), CMUN = trimws(CMUN), cusec = paste0(CPRO, CMUN))

compra21 <- compra21 |> 
  left_join(nombres_municipios, by = "cusec") |> 
  left_join(nombres_provincias, by = "CPRO") |> 
  left_join(ccaas, by = c("CPRO" = "CodigoProvincia"))

```

## Create variables per year

### 2011

```{r}

# Solve some problems in certain municipalities
propiedad <- propiedad %>%
  mutate(CPRO = case_when(
    MUN_LITERAL == "Baena" ~ "14",
    MUN_LITERAL == "Barañain" ~ "31",
    cusec == "47994" ~ "47",
    MUN_LITERAL == "Calatayud" ~ "50",
    TRUE ~ CPRO))

# Add CCAA names
compra11<-compra11 |> 
  left_join(ccaas, by = c("CPRO" = "CodigoProvincia")) 

# Tenure regime variables per municipality
compra11 <- compra11 |>
  group_by(cusec) |>
  mutate(
    Compra11Mun = round((sum(TENEN_VIV == "Compra") / n()) * 100, 2),
    Propiedad11Mun = round((sum(TENEN_VIV %in% c("Compra", "Herencia o donación")) / n()) * 100, 2),
    Herencia11Mun = round((sum(TENEN_VIV == "Herencia o donación") / n()) * 100, 2),
    Alquiler11Mun = round((sum(TENEN_VIV == "Alquiler") / n()) * 100, 2),
    Otro11Mun = round((sum(TENEN_VIV == "Otro") / n()) * 100, 2)
  ) |>
  ungroup()

# Per province
compra11 <- compra11 |>
  group_by(CPRO) |>
  mutate(
    Compra11Prov = round((sum(TENEN_VIV == "Compra") / n()) * 100, 2),
    Propiedad11Prov = round((sum(TENEN_VIV %in% c("Compra", "Herencia o donación")) / n()) * 100, 2),
    Herencia11Prov = round((sum(TENEN_VIV == "Herencia o donación") / n()) * 100, 2),
    Alquiler11Prov = round((sum(TENEN_VIV == "Alquiler") / n()) * 100, 2),
    Otro11Prov = round((sum(TENEN_VIV == "Otro") / n()) * 100, 2)
  ) |> 
  ungroup()


compra11<-compra11 |> 
  left_join(ccaas, by = c("CPRO" = "CodigoProvincia")) 

# Per CCAA
compra11 <- compra11 |>
  group_by(CCAA) |>
  mutate(
     Compra11CCAA = round((sum(TENEN_VIV == "Compra") / n()) * 100, 2),
    Propiedad11CCAA = round((sum(TENEN_VIV %in% c("Compra", "Herencia o donación")) / n()) * 100, 2),
    Herencia11CCAA = round((sum(TENEN_VIV == "Herencia o donación") / n()) * 100, 2),
    Alquiler11CCAA = round((sum(TENEN_VIV == "Alquiler") / n()) * 100, 2),
  Otro11CCAA = round((sum(TENEN_VIV == "Otro") / n()) * 100, 2)) |>
  ungroup()

```

### 2021

```{r}

# Add CCAA
compra21<-compra21 |> 
  left_join(ccaas, by = c("CPRO" = "CodigoProvincia")) 

# Per municipality
compra21 <- compra21 |> 
  group_by(cusec) |> 
  mutate(
    Propiedad21Mun = round((sum(TENEN_VIV == "Propiedad") / n()) * 100, 2),
    Alquiler21Mun = round((sum(TENEN_VIV == "Alquiler") / n()) * 100, 2),
    Otro21Mun = round((sum(TENEN_VIV == "Otro") / n()) * 100, 2)) |> 
  ungroup()

# Per province
compra21 <- compra21 |>
  group_by(CPRO) |>
  mutate(
    Propiedad21Prov = round((sum(TENEN_VIV == "Propiedad") / n()) * 100, 2),
    Alquiler21Prov = round((sum(TENEN_VIV == "Alquiler") / n()) * 100, 2),
    Otro21Prov = round((sum(TENEN_VIV == "Otro") / n()) * 100, 2)
  ) |> 
  ungroup() 

# Per CCAA
compra21 <- compra21 |> 
  mutate(CCAA = CCAA.x) |> 
  select(-CCAA.y)

compra21 <- compra21 |>
  group_by(CCAA) |>
  mutate(
    Propiedad21CCAA = round((sum(TENEN_VIV == "Propiedad") / n()) * 100, 2),
    Alquiler21CCAA = round((sum(TENEN_VIV == "Alquiler") / n()) * 100, 2),
    Otro21CCAA = round((sum(TENEN_VIV == "Otro") / n()) * 100, 2)
  ) |> 
  ungroup() 

# Join small municipalities (994)
compra21 <- compra21 |> 
  left_join(cusec_rural_2021, by = c("cusec" = "cusec_original")) |> 
  mutate(cusec = coalesce(cusec.y, cusec)) |> 
  select(-cusec.y) |> 
  group_by(cusec) |> # Group all small mun together by province 
  mutate(Propiedad21Mun = round(mean(Propiedad21Mun), 2),
            Alquiler21Mun = round(mean(Alquiler21Mun), 2),
            Otro21Mun = round(mean(Otro21Mun), 2)) |> 
  ungroup() 

#Save database with individuals of target population
#write.xlsx(compra11, "target11.xlsx")
#write.xlsx(compra21, "target21.xlsx")

```

## Create difference variables

```{r}

# Select only tenure regime variables
tenencia11 <- compra11 |> 
  select(cusec, CPRO, Compra11Mun, Propiedad11Mun, Herencia11Mun, Alquiler11Mun, Otro11Mun, Propiedad11Prov, Herencia11Prov, Alquiler11Prov, Otro11Prov, Propiedad11CCAA, Herencia11CCAA, Alquiler11CCAA, Otro11CCAA) |> 
  distinct(cusec, .keep_all = T)

tenencia21 <- compra21 |> 
  select(cusec, CPRO, Propiedad21Mun, Alquiler21Mun, Otro21Mun, Propiedad21Prov, Alquiler21Prov, Otro21Prov, Propiedad21CCAA, Alquiler21CCAA, Otro21CCAA) |> 
  distinct(cusec, .keep_all = T)

# Join both years
tenencias <- tenencia11 |> 
  left_join(tenencia21, by = "cusec")

# Create differences
tenencias <- tenencias |> 
  mutate(
    PropiedadDifMun = Propiedad21Mun - Propiedad11Mun,
    AlquilerDifMun = Alquiler21Mun - Alquiler11Mun,
    OtroDifMun = Otro21Mun - Otro11Mun,
    PropiedadDifProv = Propiedad21Prov - Propiedad11Prov,
    AlquilerDifProv = Alquiler21Prov - Alquiler11Prov,
    OtroDifProv = Otro21Prov - Otro11Prov,
    PropiedadDifCCAA = Propiedad21CCAA - Propiedad11CCAA,
    AlquilerDifCCAA = Alquiler21CCAA - Alquiler11CCAA,
    OtroDifCCAA = Otro21CCAA - Otro11CCAA)

# FINAL DATABASE
propiedad <- propiedad |> 
  left_join(tenencias, by = "cusec") |> 
  distinct(cusec, .keep_all=T) |> 
  select(-CPRO.y) |> 
  mutate(CPRO=CPRO.x)

propiedad <- propiedad |>
    filter(MUN_LITERAL != "Baena") |> # Remove because in 2021 it has more than 20k inhabitants and I don't have data from 2011
  filter(cusec != 47994) |> # No municipalities from 10k to 20k inhabitants in Valladolid 
  filter(MUN_LITERAL != "Calatayud") |>  # In 2021 it has less than 20k inhabitants
  filter(MUN_LITERAL != "Barañain") # En 2021 it has less than  20k inhabitants

# Save database
#write.xlsx(propiedad, "propiedad.xlsx")

```


